---
title: "camille_ida"
author: "Camille Karski"
date: '2022-04-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/camillekarski/Documents/Uni/DATA3888')

library(plyr)
library(tidyverse)
library(tuneR)
library(devtools)
library(ggplot2)
library(tsfeatures)
library(sf)
library(nngeo)
library(maps)
library(reshape2)
library(viridis)
library(plotly)
library(GGally)
library(class)

```

## Extracting data 

The data was downloaded from [Allen Coral Atlas](https://allencoralatlas.org/resources/) in a zip archive. Inside the unzipped file, the structure is: 

-Areaname-YYYYMMDDHHMMSS
  -Benthic-Map
    -benthic.gpkg
  -boundary
    -boundary.gpkg
  -Geomorphic-Map
    -geomorphic.gpkg (!OR .geojson)
  -license and documentation
  -stats
    -statistics.csv

We need to extract all the benthic, boundary, and geomorphic data for each area and create respective shape files. 

```{r Data extraction function}
#function for extracting features from zipped archives 
reef_f_extract = function(file_path,zip_name,feature_path,out_dir, reef_df){
  file_ls = as.character(unzip(paste0(file_path,zip_name), list = TRUE)$Name)
  feature_filename = str_subset(file_ls, feature_path)
  if (length(feature_filename) != 0){
    unzip(paste0(file_path,zip_name), files=feature_filename, exdir = out_dir)
    rugosity_df = st_read(paste0(out_dir, feature_filename))
    sf_use_s2(FALSE)
    reef_join_df = st_join(reef_df, rugosity_df)
    return(reef_join_df)
    }#end if
  return(reef_df)
} #end function
```

## Joining Reef and Geomorphic 

```{r Reading in reef data}
#reading in reef data set ! with variables as sf object using st_read
reef_df_og =st_read("./DATA3888_data/Reef_Check_with_cortad_variables_with_annual_rate_of_SST_change.csv",
                   options=c("X_POSSIBLE_NAMES=Longitude.Degrees","Y_POSSIBLE_NAMES=Latitude.Degrees"))
#setting reef_df to same CRS: WGS84 = 4326
st_crs(reef_df_og) = 4326
```

```{r Testing}
# file name containing all data: Global-Dataset-20211006223100
file_path = "./DATA3888_data/Global-Dataset-20211006223100.nosync/"
#setting new directory for where data will go 
outDir = "./DATA3888_data/geomorphic_all/"
#target file name for extraction 
geom_filepath = "^Geomorphic-Map/geomorphic"
reef_join_df = reef_df_og
test_error = "Northern-Caribbean--Florida----Bahamas-20201125220500.zip"
north_test = reef_f_extract(file_path,test_error,geom_filepath,outDir,reef_join_df)

```

```{r Data extraction}
# file name containing all data: Global-Dataset-20211006223100
file_path = "./DATA3888_data/Global-Dataset-20211006223100.nosync/"
#creating list of all files in folder 
files = list.files(path=file_path, pattern=".zip$", recursive = F)
#setting new directory for where data will go 
outDir = "./DATA3888_data/geomorphic_all/"
#target file name for extraction 
geom_filepath = "^Geomorphic-Map/geomorphic"
reef_join_df = reef_df_og

#stopped as error ocurred when trying to open 'Northern Caribbean, Florida, & Bahamas' as file over 5Gb
#removed file and continued with extraction
second_files = files[17:31]
#commented out so as not to run again 
#for (i in second_files) {
  #reef_join_df = reef_f_extract(file_path,i,geom_filepath,outDir,reef_join_df)
#} 
```

```{r Combinded classes and writing to file}
#there is an easier way to do this I am just tired rn
reef_join_c = reef_join_df %>% mutate(class = coalesce(class.x, class.y, class.x.1, class.y.1, class.x.2, class.y.2, class.x.3, class.y.3, class.x.4, class.y.4, class.x.5, class.y.5, class.x.6, class.y.6, class.x.7, class.y.7)) %>%
         select(-class.x, -class.y, -class.x.1, -class.y.1, -class.x.2, -class.y.2, -class.x.3, -class.y.3, -class.x.4, -class.y.4, -class.x.5, -class.y.5, -class.x.6, -class.y.6, -class.x.7, -class.y.7)

#write to gpkg
#st_write(reef_join_c, 'reef_geomorphic_joined.gpkg')
```

As the `Northern Caribbean, Florida, & Bahamas` region was a 5GB zipped file, we had to extract the data manually by selecting smaller areas and downloading from AllensCoralAtlas

```{r Reading in reef data with geomorphic}
#reading in reef data set ! with variables as sf object using st_read
reef_df =st_read("./DATA3888_GitHub/MARS_DATA3888_reefC4/Data/reef_geomorphic_joined.gpkg")
#setting reef_df to same CRS 
st_crs(reef_df) = 4326
```

```{r Extracting large files}
all_reef_df = reef_df
file_path = "./DATA3888_data/reef_geo_data/"
files = list.files(path=file_path, recursive = F)
for (i in files){
  rugosity_df = st_read(paste0(file_path, i))
  sf_use_s2(FALSE)
  all_reef_df = st_join(all_reef_df, rugosity_df)
}

```

```{r Combinded classes and writing to file}
#there is an easier way to do this I am just tired rn
reef_join_all = all_reef_df %>% mutate(class = coalesce(class.x, class.y, class.x.1, class.y.1, class.x.2, class.y.2, class.x.3, class.y.3)) %>%
         select(-class.x, -class.y, -class.x.1, -class.y.1, -class.x.2, -class.y.2, -class.x.3, -class.y.3)

#write to gpkg
st_write(reef_join_all, 'reef_geomorphic_joined_all.gpkg')
```

Plotting to see the differences on a map in points

```{r Quick vis}
#removing NAs for vis
reef_join_NA = reef_join_all %>% drop_na(class)
#creating df with only NAs 
only_nas = reef_join_all[is.na(reef_join_all$class),]
world_map = map_data("world")

geo_class_vis = ggplot() +  
  geom_polygon(data =world_map, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_sf(data = only_nas, aes(colour = class), alpha=1) +
  geom_sf(data = reef_join_NA, aes(colour = class), alpha=0.6) +
  theme_void()
geo_class_vis
```

The plot above still shows many NA values: 
```{r Looking at NAs}
length(reef_join_all$class) #9215
length(reef_join_NA$class) #4557
length(only_nas$class) #4658
#4557 + 4658 = 9215
#ggplotly(geo_class_vis)
```

From investigation of the data sets on the map, it looks like the reef dataset is not as precise as the Allen Coral dataset - hence points not touching some polygons. As such, to class as many reef points as possible, we will create a 200m buffer (as looking at geomorphic mapping on Allen, 200m is the smallest size that will include most datapoints). This may effect the data integrity, hence I will create a new `.gpgk` file `reef_geomorphic_joined_all_KNN.gpkg`. When modeling, we can then use both files to and calculate the difference in accuracy. 

```{r Creating buffer for NA points}
# change to Irish grid, which uses meters
only_nas_i = st_transform(only_nas, 29902)
reef_na_buffer = st_buffer(only_nas_i, 200)
#change back to 4326 to perform extraction
reef_na_buffer = st_transform(reef_na_buffer, 4326)
```

```{r Extracting class for NA values}
# file name containing all data: Global-Dataset-20211006223100
file_path = "./DATA3888_data/Global-Dataset-20211006223100.nosync/"
#creating list of all files in folder 
files = list.files(path=file_path, pattern=".zip$", recursive = F)
#setting new directory for where data will go 
outDir = "./DATA3888_data/geomorphic_all/"
#target file name for extraction 
geom_filepath = "^Geomorphic-Map/geomorphic"
reef_join_df_nas = reef_na_buffer
#commented out so as not to run again 
for (i in files) {
  reef_join_df_nas = reef_f_extract(file_path,i,geom_filepath,outDir,reef_join_df_nas)
} 
```

```{r Joining class columns}
#combining classes into 1 columns
reef_join_na_df = reef_join_df_nas %>% mutate(class = coalesce(class.x, class.y, class.x.1, class.y.1, class.x.2, class.y.2, class.x.3, class.y.3, class.x.4, class.y.4, class.x.5, class.y.5, class.x.6, class.y.6, class.x.7, class.y.7, class.x.8, class.y.8, class.x.9, class.y.9, class.x.10, class.y.10, class.x.11, class.y.11, class.x.12, class.y.12, class.x.13, class.y.13, class.x.14, class.y.14)) %>%
         select(-class.x, -class.y, -class.x.1, -class.y.1, -class.x.2, -class.y.2, -class.x.3, -class.y.3, -class.x.4, -class.y.4, -class.x.5, -class.y.5, -class.x.6, -class.y.6, -class.x.7, -class.y.7,-class.x.8, -class.y.8, -class.x.9, -class.y.9, -class.x.10, -class.y.10, -class.x.11, -class.y.11, -class.x.12, -class.y.12, -class.x.13, -class.y.13, -class.x.14, -class.y.14)

```

```{r Extracting large NA data}
all_reef_df_na = reef_join_na_df
file_path = "./DATA3888_data/reef_geo_data/"
files = list.files(path=file_path, recursive = F)
for (i in files){
  rugosity_df = st_read(paste0(file_path, i))
  sf_use_s2(FALSE)
  all_reef_df_na = st_join(all_reef_df_na, rugosity_df)
}

```

```{r Joining class columns}
#there is an easier way to do this I am just tired rn
all_reef_df_na = all_reef_df_na %>% mutate(class_na = coalesce(class.x, class.y, class.x.1, class.y.1, class.x.2, class.y.2, class)) %>%
         select(-class.x, -class.y, -class.x.1, -class.y.1, -class.x.2, -class.y.2, -class)
#Simple feature collection with 10290 features and 56 fields

#write to gpkg 
#st_write(all_reef_df_na, 'reef_geomorphic_joined_nas.gpkg')
```

```{r Looking at NAs}
only_nas = all_reef_df_na[is.na(all_reef_df_na$class_na),]
length(reef_join_all$class) #9215
length(all_reef_df_na$class_na) #10290
length(only_nas$class) #2160
```

```{r Removing NAs and selecting distinct rows}
#all_reef_df_c = all_reef_df_na %>% drop_na()
#all_reef_df_c = distinct(all_reef_df_c)
#st_write(all_reef_df_c, 'reef_geomorphic_joined_nas_c.gpkg')
```

```{r}
table(rownames(all_reef_df_c))
```


```{r}
reef_joined_na_c = st_read('./DATA3888_GitHub//MARS_DATA3888_reefC4/Data/reef_geomorphic_joined_nas_c.gpkg')
```
```{r}
reef_joined_na_c
```

```{r}
reef_joined_all = st_read('./DATA3888_GitHub//MARS_DATA3888_reefC4/Data/reef_geomorphic_joined_all.gpkg')
```

```{r}
reef_joined_all = reef_joined_all %>% drop_na(class)
```

## Figures and Shiny App

Correlation plot - can use rugosity, diversity, SSTA_std, and average bleaching. 
```{r Q4 Customizing correlation plot}
#Customizing correlation plot with colours 
#code adapted from https://github.com/ggobi/ggally/issues/139

my_custom_cor_color <- function(data, mapping, color = I("black"), sizeRange = c(1, 5), ...) {

  #eval(mapping$x, data) should be switched to GGally::eval_data_col(data, mapping$x)
  # get the x and y data to use the other code
  x <- GGally::eval_data_col(data, mapping$x)
  y <- GGally::eval_data_col(data, mapping$y)

  ct <- cor.test(x,y)
  
  sig <- symnum(ct$p.value, corr = FALSE, na = FALSE,
                cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                symbols = c("***", "**", "*", ".", " "))

  r <- unname(ct$estimate)
  rt <- format(r, digits=2)[1]
  tt <- as.character(rt)

  # plot the cor value
  p <- ggally_text(
   label = tt, 
   mapping = aes(),
   xP = 0.5, yP = 0.5, 
   size = 6,
   color=color,
   ...
  ) + 
    # add the sig stars
    geom_text(aes_string(x = 0.7, y = 0.7),
      label = sig, 
      color = color,
      ...
    ) +
    geom_text(aes_string(x = 0.4, y = 0.7),
              label = "Corr: ",
              color = color
              ) + 
    # remove all the background stuff and wrap it with a dashed line
    theme(
      panel.background=element_rect(fill="white", linetype = "dashed"), 
      panel.grid.minor=element_blank(), 
      panel.grid.major=element_blank()) 

  corColors <- RColorBrewer::brewer.pal(n = 7, name = "PuOr") #[2:6]

  if (r <= -0.08) {
    corCol <- corColors[7]
  } else if (r <= -0.06) {
    corCol <- corColors[6]
  } else if (r < 0.06) {
    corCol <- corColors[5]
  } else if (r < 0.08) {
    corCol <- corColors[3]
  } else {
    corCol <- corColors[2]
  }
  p <- p + theme(
    panel.background = element_rect(fill= corCol)
  )

  p
}

my_custom_smooth <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_point(color = I("orange")) + 
    scale_color_brewer(palette="PuOr") +
    geom_smooth(method = "lm", color = I("black")) + 
    theme(panel.background=element_blank(),
        panel.grid.minor=element_blank(),
        panel.grid.major=element_line(color="lightgray"))
}

```
```{r}
reef_joined_corri = reef_joined_all %>% select(Reef.ID, Average_bleaching, SSTA_Frequency_Standard_Deviation, Diversity, rate_of_SST_change, class)
reef_joined_corri = data.frame(reef_joined_corri)
```
```{r}
reef_joined_corri = transform(reef_joined_corri, Average_bleaching = as.numeric(Average_bleaching), 
               SSTA_Frequency_Standard_Deviation = as.numeric(SSTA_Frequency_Standard_Deviation),
          Diversity = as.numeric(Diversity), rate_of_SST_change = as.numeric(rate_of_SST_change))
```


```{r Q4 Correlation plot, fig.cap= "Fig 3.3: Variable Correlation Visualisation", message=FALSE, warning=FALSE}
reef_corri = ggpairs(reef_joined_corri, columns=2:6,
        title = "Correlation of Bleaching and other variables",
        columnLabels = c("Average_bleaching","SSTA_Frequency_Standard_Deviation",
                         "Diversity", "rate_of_SST_change", "class"),
        upper=list(continuous=wrap(my_custom_cor_color)),
        cardinality_threshold=NULL,
        lower=list(continuous=wrap(my_custom_smooth), 
                   combo = wrap("dot", color=Class)))         
reef_corri
```
```{r}
ggpairs(reef_joined_corri, columns=2:6, 
        columnLabels =c("Average_bleaching","SSTA_Frequency_Standard_Deviation",
                         "Diversity", "rate_of_SST_change", "class"),
        cardinality_threshold=NULL)
```



```{r}
#reef_joined_all
pca_eMat = prcomp(reef_joined_all, scale = TRUE)
options(digits = 2)
summary(pca_eMat)
```










