---
title: "camille_ida"
author: "Camille Karski"
date: '2022-04-09'
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/camillekarski/Documents/Uni/DATA3888/')

library(plyr)
library(tidyverse)
library(tuneR)
library(devtools)
library(ggplot2)
library(tsfeatures)
library(sf)
library(nngeo)
library(maps)
library(reshape2)
library(viridis)
library(plotly)
library(GGally)
library(class)

```

## Extracting data 

The data was downloaded from [Allen Coral Atlas](https://allencoralatlas.org/resources/) in a zip archive. Inside the unzipped file, the structure is: 

-Areaname-YYYYMMDDHHMMSS
  -Benthic-Map
    -benthic.gpkg
  -boundary
    -boundary.gpkg
  -Geomorphic-Map
    -geomorphic.gpkg (!OR .geojson)
  -license and documentation
  -stats
    -statistics.csv

We need to extract all the benthic, boundary, and geomorphic data for each area and create respective shape files. 

```{r Data extraction function}
#function for extracting features from zipped archives 
reef_f_extract = function(file_path,zip_name,feature_path,out_dir, reef_df){
  file_ls = as.character(unzip(paste0(file_path,zip_name), list = TRUE)$Name)
  feature_filename = str_subset(file_ls, feature_path)
  if (length(feature_filename) != 0){
    unzip(paste0(file_path,zip_name), files=feature_filename, exdir = out_dir)
    rugosity_df = st_read(paste0(out_dir, feature_filename))
    sf_use_s2(FALSE)
    reef_join_df = st_join(reef_df, rugosity_df)
    return(reef_join_df)
    }#end if
  return(reef_df)
} #end function
```

## Joining Reef and Geomorphic 

```{r Reading in reef data}
#reading in reef data set ! with variables as sf object using st_read
reef_df_og =st_read("./DATA3888_data/Reef_Check_with_cortad_variables_with_annual_rate_of_SST_change.csv",
                   options=c("X_POSSIBLE_NAMES=Longitude.Degrees","Y_POSSIBLE_NAMES=Latitude.Degrees"))
#setting reef_df to same CRS: WGS84 = 4326
st_crs(reef_df_og) = 4326
```

```{r Testing}
# file name containing all data: Global-Dataset-20211006223100
file_path = "./DATA3888_data/Global-Dataset-20211006223100.nosync/"
#setting new directory for where data will go 
outDir = "./DATA3888_data/geomorphic_all/"
#target file name for extraction 
geom_filepath = "^Geomorphic-Map/geomorphic"
reef_join_df = reef_df_og
test_error = "Northern-Caribbean--Florida----Bahamas-20201125220500.zip"
north_test = reef_f_extract(file_path,test_error,geom_filepath,outDir,reef_join_df)

```

```{r Data extraction}
# file name containing all data: Global-Dataset-20211006223100
file_path = "./DATA3888_data/Global-Dataset-20211006223100.nosync/"
#creating list of all files in folder 
files = list.files(path=file_path, pattern=".zip$", recursive = F)
#setting new directory for where data will go 
outDir = "./DATA3888_data/geomorphic_all/"
#target file name for extraction 
geom_filepath = "^Geomorphic-Map/geomorphic"
reef_join_df = reef_df_og

#stopped as error ocurred when trying to open 'Northern Caribbean, Florida, & Bahamas' as file over 5Gb
#removed file and continued with extraction
second_files = files[17:31]
#commented out so as not to run again 
#for (i in second_files) {
  #reef_join_df = reef_f_extract(file_path,i,geom_filepath,outDir,reef_join_df)
#} 
```

```{r Combinded classes and writing to file}
#there is an easier way to do this I am just tired rn
reef_join_c = reef_join_df %>% mutate(class = coalesce(class.x, class.y, class.x.1, class.y.1, class.x.2, class.y.2, class.x.3, class.y.3, class.x.4, class.y.4, class.x.5, class.y.5, class.x.6, class.y.6, class.x.7, class.y.7)) %>%
         select(-class.x, -class.y, -class.x.1, -class.y.1, -class.x.2, -class.y.2, -class.x.3, -class.y.3, -class.x.4, -class.y.4, -class.x.5, -class.y.5, -class.x.6, -class.y.6, -class.x.7, -class.y.7)

#write to gpkg
#st_write(reef_join_c, 'reef_geomorphic_joined.gpkg')
```

As the `Northern Caribbean, Florida, & Bahamas` region was a 5GB zipped file, we had to extract the data manually by selecting smaller areas and downloading from AllensCoralAtlas

```{r Reading in reef data with geomorphic}
#reading in reef data set ! with variables as sf object using st_read
reef_df =st_read("./DATA3888_GitHub/MARS_DATA3888_reefC4/Data/reef_geomorphic_joined.gpkg")
#setting reef_df to same CRS 
st_crs(reef_df) = 4326
```

```{r Extracting large files}
all_reef_df = reef_df
file_path = "./DATA3888_data/reef_geo_data/"
files = list.files(path=file_path, recursive = F)
for (i in files){
  rugosity_df = st_read(paste0(file_path, i))
  sf_use_s2(FALSE)
  all_reef_df = st_join(all_reef_df, rugosity_df)
}

```

```{r Combinded classes and writing to file}
#there is an easier way to do this I am just tired rn
reef_join_all = all_reef_df %>% mutate(class = coalesce(class.x, class.y, class.x.1, class.y.1, class.x.2, class.y.2, class.x.3, class.y.3)) %>%
         select(-class.x, -class.y, -class.x.1, -class.y.1, -class.x.2, -class.y.2, -class.x.3, -class.y.3)

#write to gpkg
st_write(reef_join_all, 'reef_geomorphic_joined_all.gpkg')
```

Plotting to see the differences on a map in points

```{r Quick vis}
#removing NAs for vis
reef_join_NA = reef_join_all %>% drop_na(class)
#creating df with only NAs 
only_nas = reef_join_all[is.na(reef_join_all$class),]
world_map = map_data("world")

geo_class_vis = ggplot() +  
  geom_polygon(data =world_map, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_sf(data = only_nas, aes(colour = class), alpha=1) +
  geom_sf(data = reef_join_NA, aes(colour = class), alpha=0.6) +
  theme_void()
geo_class_vis
```

The plot above still shows many NA values: 
```{r Looking at NAs}
length(reef_join_all$class) #9215
length(reef_join_NA$class) #4557
length(only_nas$class) #4658
#4557 + 4658 = 9215
#ggplotly(geo_class_vis)
```

From investigation of the data sets on the map, it looks like the reef dataset is not as precise as the Allen Coral dataset - hence points not touching some polygons. As such, to class as many reef points as possible, we will create a 200m buffer (as looking at geomorphic mapping on Allen, 200m is the smallest size that will include most datapoints). This may effect the data integrity, hence I will create a new `.gpgk` file `reef_geomorphic_joined_all_KNN.gpkg`. When modeling, we can then use both files to and calculate the difference in accuracy. 

```{r Creating buffer for NA points}
# change to Irish grid, which uses meters
only_nas_i = st_transform(only_nas, 29902)
reef_na_buffer = st_buffer(only_nas_i, 200)
#change back to 4326 to perform extraction
reef_na_buffer = st_transform(reef_na_buffer, 4326)
```

```{r Extracting class for NA values}
# file name containing all data: Global-Dataset-20211006223100
file_path = "./DATA3888_data/Global-Dataset-20211006223100.nosync/"
#creating list of all files in folder 
files = list.files(path=file_path, pattern=".zip$", recursive = F)
#setting new directory for where data will go 
outDir = "./DATA3888_data/geomorphic_all/"
#target file name for extraction 
geom_filepath = "^Geomorphic-Map/geomorphic"
reef_join_df_nas = reef_na_buffer
#commented out so as not to run again 
for (i in files) {
  reef_join_df_nas = reef_f_extract(file_path,i,geom_filepath,outDir,reef_join_df_nas)
} 
```

```{r Joining class columns}
#combining classes into 1 columns
reef_join_na_df = reef_join_df_nas %>% mutate(class = coalesce(class.x, class.y, class.x.1, class.y.1, class.x.2, class.y.2, class.x.3, class.y.3, class.x.4, class.y.4, class.x.5, class.y.5, class.x.6, class.y.6, class.x.7, class.y.7, class.x.8, class.y.8, class.x.9, class.y.9, class.x.10, class.y.10, class.x.11, class.y.11, class.x.12, class.y.12, class.x.13, class.y.13, class.x.14, class.y.14)) %>%
         select(-class.x, -class.y, -class.x.1, -class.y.1, -class.x.2, -class.y.2, -class.x.3, -class.y.3, -class.x.4, -class.y.4, -class.x.5, -class.y.5, -class.x.6, -class.y.6, -class.x.7, -class.y.7,-class.x.8, -class.y.8, -class.x.9, -class.y.9, -class.x.10, -class.y.10, -class.x.11, -class.y.11, -class.x.12, -class.y.12, -class.x.13, -class.y.13, -class.x.14, -class.y.14)

```

```{r Extracting large NA data}
all_reef_df_na = reef_join_na_df
file_path = "./DATA3888_data/reef_geo_data/"
files = list.files(path=file_path, recursive = F)
for (i in files){
  rugosity_df = st_read(paste0(file_path, i))
  sf_use_s2(FALSE)
  all_reef_df_na = st_join(all_reef_df_na, rugosity_df)
}

```

```{r Joining class columns}
#there is an easier way to do this I am just tired rn
all_reef_df_na = all_reef_df_na %>% mutate(class_na = coalesce(class.x, class.y, class.x.1, class.y.1, class.x.2, class.y.2, class)) %>%
         select(-class.x, -class.y, -class.x.1, -class.y.1, -class.x.2, -class.y.2, -class)
#Simple feature collection with 10290 features and 56 fields

#write to gpkg 
#st_write(all_reef_df_na, 'reef_geomorphic_joined_nas.gpkg')
```

```{r Looking at NAs}
only_nas = all_reef_df_na[is.na(all_reef_df_na$class_na),]
length(reef_join_all$class) #9215
length(all_reef_df_na$class_na) #10290
length(only_nas$class) #2160
```

```{r Removing NAs and selecting distinct rows}
#all_reef_df_c = all_reef_df_na %>% drop_na()
#all_reef_df_c = distinct(all_reef_df_c)
#st_write(all_reef_df_c, 'reef_geomorphic_joined_nas_c.gpkg')
```

As there were too many double ups of classes for the same points, we decided to use only the data we had an exact match to, as an incorrect geomorphic type could change the rugosity estimate. 

## Figures 
```{r Reading in data}
reef_geomorphic = st_read("./DATA3888_Github/MARS_DATA3888_reefC4/Data/reef_geomorphic_joined_all.gpkg")
reef_geomorphic = reef_geomorphic %>% 
  mutate(Date = as.Date(Date, "%d-%b-%y")) %>%
  mutate(Depth = as.numeric(Depth)) %>%
  filter(Depth <= 15) # as per Allen Coral Atlas specs

reef_recent = reef_geomorphic %>%
  select(Reef.ID, Reef.Name, Country, Date, SSTA_Frequency_Standard_Deviation, rate_of_SST_change, Depth, Diversity, class,Average_bleaching, geom) %>% 
  mutate(SSTA_Frequency_Standard_Deviation = as.numeric(SSTA_Frequency_Standard_Deviation),
         rate_of_SST_change = as.numeric(rate_of_SST_change),
         Average_bleaching = as.numeric(Average_bleaching),
         Depth = as.numeric(Depth), 
         Diversity = as.numeric(Diversity)) %>%
  group_by(Reef.ID) %>% filter(Date == max(Date))

#bleached or not bleached
reef_recent["bleached"] = ifelse(reef_recent$Average_bleaching > 0, 1,0)

#Rugosity
reef_recent["rugosity"] = ifelse(reef_recent$class == "Reef Slope" | reef_recent$class == "Outer Reef Flat","High", ifelse(reef_recent$class == "Sheltered Reef Slope" | reef_recent$class == "Reef Crest" | reef_recent$class == "Back Reef Slope","Medium","Low"))

#Rugosity
reef_recent["rugosity_num"] = ifelse(reef_recent$class == "Reef Slope" | reef_recent$class == "Outer Reef Flat",3, ifelse(reef_recent$class == "Sheltered Reef Slope" | reef_recent$class == "Reef Crest" | reef_recent$class == "Back Reef Slope",2,1))
#Remove NA
reef_final = reef_recent %>% drop_na()

#st_write(reef_final, 'reef_rugosity_cleaned.gpkg')
```

Correlation plot - can use rugosity, diversity, SSTA_std, and average bleaching. 

```{r Correlation plot df}
reef_joined_corri = reef_final %>% select(Reef.ID, Average_bleaching, SSTA_Frequency_Standard_Deviation, Diversity, rate_of_SST_change, rugosity_num, class)
reef_joined_corri = data.frame(reef_joined_corri)
```

```{r Customizing correlation plot}
#Customizing correlation plot with colours 
#code adapted from https://github.com/ggobi/ggally/issues/139

my_custom_cor_color <- function(data, mapping, color = I("black"), sizeRange = c(1, 5), ...) {

  #eval(mapping$x, data) should be switched to GGally::eval_data_col(data, mapping$x)
  # get the x and y data to use the other code
  x <- GGally::eval_data_col(data, mapping$x)
  y <- GGally::eval_data_col(data, mapping$y)

  ct <- cor.test(x,y)
  
  sig <- symnum(ct$p.value, corr = FALSE, na = FALSE,
                cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                symbols = c("***", "**", "*", ".", " "))

  r <- unname(ct$estimate)
  rt <- format(r, digits=2)[1]
  tt <- as.character(rt)

  # plot the cor value x and y for number 
  p <- ggally_text(
   label = tt, 
   mapping = aes(),
   xP = 0.5, yP = 0.5, 
   size = 4,
   color=color,
   ...
  ) + 
    # add the sig stars #x and y are position of stars
    geom_text(aes_string(x = 0.7, y = 0.75),
      label = sig, 
      color = color,
      ...
    ) +   #x and y are position or text Corr: 
    geom_text(aes_string(x = 0.5, y = 0.8),
              size = 3,
              label = "Corr: ",
              color = color
              ) + 
    # remove all the background stuff and wrap it with a dashed line
    theme(
      panel.background=element_rect(fill="white", linetype = "dashed"), 
      panel.grid.minor=element_blank(), 
      panel.grid.major=element_blank()) 

  #corColors <- RColorBrewer::brewer.pal(n = 7, name = "RdBu") #[2:6]
  #"#B35806" "#E08214" "#D1E5F0" "#4393C3" "#2166AC" 

  if (r <= -0.08) {
    corCol <- "#2166AC"
  } else if (r <= -0.06) {
    corCol <- "#4393C3"
  } else if (r < 0.06) {
    corCol <- "#D1E5F0"
  } else if (r < 0.08) {
    corCol <- "#E08214"
  } else {
    corCol <- "#B35806"
  }
  p <- p + theme(
    panel.background = element_rect(fill= corCol)
  )

  p
}

my_custom_smooth <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_point(color = I("orange"), alpha =0.2) + 
    scale_color_brewer(palette="RdBu") +
    geom_smooth(method = "lm", color = I("black")) + 
    theme(panel.background=element_blank(),
        panel.grid.minor=element_blank(),
        panel.grid.major=element_line(color="lightgray"))
}

```

```{r Correlation plot, fig.cap= "Fig 3.3: Variable Correlation Visualisation", message=FALSE, warning=FALSE}
reef_corri = ggpairs(reef_joined_corri, columns=2:7,
        title = "Correlation of Bleaching and chosen variables",
        columnLabels = c("Averageb Bleaching","SSTA Frequency STD",
                         "Diversity", "Rate of SST change", "Rugosity", "Class"),
        upper=list(continuous=my_custom_cor_color),
        cardinality_threshold=NULL,
        lower=list(continuous=my_custom_smooth), 
                   combo = "dot", color=Class)         
reef_corri
```

```{r Saving final reef file for app}
reef_geomorphic = st_read("/Users/camillekarski/Documents/Uni/DATA3888/DATA3888_GitHub/MARS_DATA3888_reefC4/Data/reef_geomorphic_joined_all.gpkg")

#adding to column with new long co-ords for world2 map
wrapfunction = function(x){ifelse(x > 0, x, 360+x)}
reef_geomorphic$wrapLongitute = sapply(reef_geomorphic$Longitude.Degrees, wrapfunction)

reef_geomorphic = reef_geomorphic %>% 
  #as.data.frame() %>%
  mutate(Date = as.Date(Date, "%d-%b-%y")) %>%
  mutate(Date = as.numeric(format(Date, "%Y"))) %>%
  mutate(Depth = as.numeric(Depth)) %>%
  filter(Depth <= 15) # as per Allen Coral Atlas specs

reef_recent = reef_geomorphic %>%
  select(Reef.ID, Reef.Name, Country, Date, SSTA_Frequency_Standard_Deviation, Depth, Diversity, class, Average_bleaching, Latitude.Degrees, wrapLongitute) %>% 
  mutate(SSTA_Frequency_Standard_Deviation = as.numeric(SSTA_Frequency_Standard_Deviation),
         Average_bleaching = as.numeric(Average_bleaching),
         Depth = as.numeric(Depth), 
         Diversity = as.numeric(Diversity)) %>%
  group_by(Reef.ID)

reef_recent["rugosity"] = ifelse(reef_recent$class == "Reef Slope" | reef_recent$class == "Outer Reef Flat","High", ifelse(reef_recent$class == "Sheltered Reef Slope" | reef_recent$class == "Reef Crest" | reef_recent$class == "Back Reef Slope","Medium","Low"))
reef_recent["rugosity_num"] = ifelse(reef_recent$class == "Reef Slope" | reef_recent$class == "Outer Reef Flat",3, ifelse(reef_recent$class == "Sheltered Reef Slope" | reef_recent$class == "Reef Crest" | reef_recent$class == "Back Reef Slope",2,1))

reef_recent["bleached"] = ifelse(reef_recent$Average_bleaching > 0,1,0)

reef_final = reef_recent %>% 
  drop_na() %>% 
  mutate(class = as.factor(class)) %>%
  mutate(rugosity = factor(rugosity, levels = c("Low", "Medium", "High")))
#st_write(reef_final, 'reef_final.gpkg')
```

          