---
title: "nathan_ida"
author: '490418846'
date: '2022-04-02'
output: html_document
---

```{r}
library(tidyverse)
library(sf)
library(ggplot2)
library(maps)
library(viridis)
```

## Reef data

Reading it in and doing some cleaning

```{r}
reef_df = read.csv("Data/Reef_Check_with_cortad_variables_with_annual_rate_of_SST_change.csv")
reef_df = reef_df %>% 
  drop_na(Longitude.Degrees) %>%
  drop_na(Latitude.Degrees) %>%
  mutate(Date = as.Date(Date, "%d-%b-%y")) %>%
  mutate(Average_bleaching = as.numeric(Average_bleaching))

reef_df$geom = reef_df %>% st_as_sf(coords = c("Longitude.Degrees", "Latitude.Degrees"),crs=4326)
```

## Merged dataset

Merging with geomorphic data and then cleaning

```{r}
reef_geomorphic = st_read("Data/reef_geomorphic_joined_all.gpkg")
reef_geomorphic = reef_geomorphic %>% 
  as.data.frame() %>%
  mutate(Date = as.Date(Date, "%d-%b-%y")) %>%
  mutate(Depth = as.numeric(Depth)) %>%
  filter(Depth <= 15) # as per Allen Coral Atlas specs

reef_geomorphic
```

## Investigating NAs

```{r}
not_na = reef_geomorphic %>% 
  slice(which(!is.na(reef_geomorphic$class))) %>%
  select(Reef.ID, Depth, Average_bleaching, class) %>%
  as.data.frame()

not_na # 4545 - still some missing data as we haven't finalised which dataset to use

not_na %>% distinct(Reef.ID, Depth, class) # 3036 - meaning there are some duplicates

not_na %>% distinct(Reef.ID, Depth) # 3036 - meaning that the same reefs but at different depths have the same class identifier, which might be a problem
```

## Frequency of bleaching

Taken by how many dated entries there are for each reef, ignoring depth (as they are usually taken on the same day).

```{r}
reef_freq = reef_geomorphic %>%
  select(Reef.ID, Date, Longitude.Degrees, Latitude.Degrees) %>%
  distinct(Reef.ID, Longitude.Degrees, Latitude.Degrees, Date) %>%
  group_by(Reef.ID, Longitude.Degrees, Latitude.Degrees) %>%
  summarise(count = n())

reef_freq

world_map = map_data("world")
ggplot() + 
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = "grey", alpha = 0.3) +
  geom_point(data = reef_freq, alpha = 0.2, aes(y = Latitude.Degrees, x = Longitude.Degrees, size = count, colour = count)) + 
  scale_colour_viridis() + 
  theme_minimal()
```

## Severity of bleaching

Taking the difference between the bleaching data of the most recent entry and the entry before that for each reef, if data exists (?). Will take the average of bleaching across different depths.

```{r}
# Bleaching record of the most recent date
reef_sev1 = reef_geomorphic %>%
  select(Reef.ID, Date, Average_bleaching, Longitude.Degrees, Latitude.Degrees) %>% 
  mutate(Average_bleaching = as.numeric(Average_bleaching)) %>%
  group_by(Reef.ID, Date, Longitude.Degrees, Latitude.Degrees) %>%
  summarise(bleaching = mean(Average_bleaching)) %>%
  group_by(Reef.ID) %>%
  filter(Date == max(Date))

# Bleaching record of the second most recent date
reef_sev2 = reef_geomorphic %>%
  select(Reef.ID, Date, Average_bleaching, Longitude.Degrees, Latitude.Degrees) %>% 
  mutate(Average_bleaching = as.numeric(Average_bleaching)) %>%
  group_by(Reef.ID, Date, Longitude.Degrees, Latitude.Degrees) %>%
  summarise(bleaching = mean(Average_bleaching)) %>%
  group_by(Reef.ID) %>%
  filter(Date == sort(Date, decreasing = TRUE)[2])

# Merging and taking the difference between them
reef_sev = merge(reef_sev1, reef_sev2, by = "Reef.ID") %>%
  mutate(Average_bleaching = bleaching.x - bleaching.y) %>%
  select(Reef.ID, Longitude.Degrees.x, Latitude.Degrees.x, Average_bleaching) %>%
  rename(Longitude.Degrees = Longitude.Degrees.x) %>%
  rename(Latitude.Degrees = Latitude.Degrees.x)

reef_sev

ggplot() + 
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = "grey", alpha = 0.3) +
  geom_point(data = reef_sev, alpha = 0.2, aes(y = Latitude.Degrees, x = Longitude.Degrees, size = Average_bleaching, colour = Average_bleaching)) + 
  scale_colour_viridis() + 
  theme_minimal()
```


## Deprecated code

```{r}
# storms_list = read.csv("Data/ibtracs.ALL.list.v04r00.csv")
# storms_list = storms_list %>% 
#   slice(-1) %>%
#   select(SID, SEASON, LAT, LON, STORM_SPEED, STORM_DIR) %>%
#   mutate(SEASON = as.numeric(SEASON)) %>%
#   filter(SEASON >= 1997) %>%
#   mutate(LAT = as.numeric(LAT)) %>%
#   mutate(LON = as.numeric(LON)) %>%
#   mutate(STORM_SPEED = as.numeric(STORM_SPEED)) %>%
#   mutate(STORM_DIR = as.numeric(STORM_DIR)) %>%
#   drop_na()
# storms_list
```

```{r}
# benthic = st_read("Data/benthic.gpkg")
# geom = lapply(benthic$geom, st_bbox)
# bbox_vector = c()
# for (i in 1:length(geom)) {
#   bbox_vector = rbind(bbox_vector, t(matrix(unlist(geom[[i]]))))
# }
# benthic = as.data.frame(cbind(benthic$class, bbox_vector))
# colnames(benthic) = c("Class", "xmin", "ymin", "xmax", "ymax")
# benthic = benthic %>%
#   mutate(xmin = as.numeric(xmin)) %>%
#   mutate(ymin = as.numeric(ymin)) %>%
#   mutate(xmax = as.numeric(xmax)) %>%
#   mutate(ymax = as.numeric(ymax))
# benthic
```

```{r}
# geomorphic = st_read("Data/geomorphic.gpkg")
# geom = lapply(geomorphic$geom, st_bbox)
# bbox_vector = c()
# for (i in 1:length(geom)) {
#   bbox_vector = rbind(bbox_vector, t(matrix(unlist(geom[[i]]))))
# }
# geomorphic = as.data.frame(cbind(geomorphic$class, bbox_vector))
# colnames(geomorphic) = c("Class", "xmin", "ymin", "xmax", "ymax")
# geomorphic = geomorphic %>%
#   mutate(xmin = as.numeric(xmin)) %>%
#   mutate(ymin = as.numeric(ymin)) %>%
#   mutate(xmax = as.numeric(xmax)) %>%
#   mutate(ymax = as.numeric(ymax))
# geomorphic
```
```

